---
title: "Good Software Engineering Practice for R Packages"
---

<!--
Creators (initial authors): 
- Daniel Sabanes Bove, https://github.com/danielinteractive/ and www.linkedin.com/in/danielsabanesbove/
- Friedrich Pahlke, https://github.com/fpahlke/ and www.linkedin.com/in/pahlke/
- Kevin Kunzmann, https://github.com/kkmann/ and https://www.linkedin.com/in/kevin-kunzmann-6486a11bb/
- Liming Li, https://github.com/clarkliming
- Joe Zhu, https://github.com/shajoezhu and http://www.linkedin.com/in/joe-zhu-464b5818
- Shuang Li, https://github.com/shuangli22
In the current version, changes were done by (later authors):
- Doug Kelkhoff, https://github.com/dgkf and https://www.linkedin.com/in/doug-kelkhoff/
- Philippe Boileau, https://github.com/philboileau and https://www.linkedin.com/in/philippe-boileau-773270205/
License: This work is licensed under the Creative Commons Attribution-ShareAlike 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by-sa/4.0/.
The source files are hosted at https://github.com/openpharma/workshop-r-swe-mtl, which is forked from
the original version at https://github.com/openpharma/workshop-r-swe.
Important: To use this work you must provide the name of the creators (initial authors), a link to the material, a link to the license, and indicate if changes were made.
--> 

Welcome to the homepage of the workshop *"Good Software Engineering Practice for R Packages"*. In this course participants will learn hands-on skills and tools to engineer reliable R packages used in biostatistics. The workshop will be conducted in two afternoons and will be a mix of presentations and exercises. Participants need to be comfortable with writing functions in R and use their own laptops as a prerequisite.

## Event Details

The event will be held on Monday 16th and Tuesday 17th October 2023 as an in-person workshop at McGill University, Montreal, Canada.
Please register [here](https://tinyurl.com/gswep4r-mtl23) by September 15th 2023.

This event is organized by the McGill Initiative in Computational Medicine and the ASA BIOP Software Engineering Working Group (SWE WG). We are grateful to have received logistic support by McGill University, in particular Adrien Osakwe and Larisa Morales Soto. The presenters in this workshop will be Doug Kelkhoff, Philippe Boileau and Daniel Sabanes Bove.

## Communication

We offer a [gitter chat channel](https://matrix.to/#/#workshop-r-swe_community:gitter.im) to communicate before, during, and after the course.

## Workshop Program

### Day 1: 16th October

```{r}
#| echo: false
readr::read_csv("slides/resources/program.csv", col_types = "cc") |>
  dplyr::filter(day == "1") |> 
  dplyr::select(- day) |> 
  knitr::kable(col.names = c("", ""))
```

### Day 2: 17th October

```{r}
#| echo: false
readr::read_csv("slides/resources/program.csv", col_types = "cc") |>
  dplyr::filter(day == "2") |>
  dplyr::select(- day) |> 
  knitr::kable(col.names = c("", ""))
```

## Prerequisites & Technical Setup

Prior to the course, participants should

-   set up a (free) [GitHub.com account](https://github.com/join). There are other git Platforms like [Gitlab](https://about.gitlab.com/) or [Bitbucket](https://bitbucket.org/) but we made the choice to go with GitHub.com for the course since it is by far the most relevant git platform in the R community.
-   download and extract [simulatr.zip](slides/download/simulatr.zip)
-   either make sure they have a working R software development setup on their own laptop (up-to-date git/Rtools/R/RStudio) or get access to <https://posit.cloud>. The latter offers a free tier account with 25 hours of computing time per months and can be accessed using ones GitHub.com account.

For the course, participants are required to use their own laptop to be able to participate in the exercises.

### Optional reading list

-   Excellent and very comprehensive [R Packages (2e)](https://r-pkgs.org/)
-   (Hardcore) description of how [Writing R Extensions](https://cran.r-project.org/doc/manuals/R-exts.html) works
-   [GitHub 'Hello World'](https://docs.github.com/en/get-started/quickstart/hello-world) tutorial on how to use GitHub.com (does not require git command line knowledge)
-   Tutorial on [version control with git and SVN](https://support.posit.co/hc/en-us/articles/200532077-Version-Control-with-Git-and-SVN) (we will be using git)
-   Command line git mini intro [trygit](https://trygit.js.org/)
-   [first steps with git & github by RStudio](https://www.youtube.com/watch?v=jN6tvgt3GK8)
-   [RStudio cheatsheets/ Git & GitHub](https://github.com/rstudio/cheatsheets/blob/main/git-github.pdf)

## Past Events

- [10th February 2023, Basel, CH](https://kkmann.github.io/workshop-r-swe/)
- [24th March 2023, Shanghai, CN](https://openpharma.github.io/workshop-r-swe/)
- [20th July 2023, San José, CA, US](https://openpharma.github.io/workshop-r-swe-sf/)
- [26th September 2023, Rockville, MD, US](https://openpharma.github.io/workshop-r-swe-md/)
